formula1

### 2020-06-01 inst/f1points.Rmd
incomplete argument to ordertable2supp()

\small{
\begin{CodeChunk}
\begin{CodeOutput}
           AUS CHN BHR RUS ESP MON CAN AZE AUT GBR HUN BEL ITA SIN MAL JPN USA MEX BRA ABU
Hamilton     2   1   2   4   1   7   1   5   4   1   4   1   1   1   2   1   1   9   4   2
Vettel       1   2   1   2   2   1   4   4   2   7   1   2   3 Ret   4 Ret   2   4   1   3
Bottas       3   6   3   1 Ret   4   2   2   1   2   3   5   2   3   5   4   5   2   2   1
Raikkonen    4   5   4   3 Ret   2   7  14   5   3   2   4   5 Ret DNS   5   3   3   3   4
Ricciardo  Ret   4   5 Ret   3   3   3   1   3   5 Ret   3   4   2   3   3 Ret Ret   6 Ret
...
Hartley      0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0  13 Ret Ret  15
Button       0   0   0   0   0 Ret   0   0   0   0   0   0   0   0   0   0   0   0   0   0
Resta        0   0   0   0   0   0   0   0   0   0 Ret   0   0   0   0   0   0   0   0   0
\end{CodeOutput}
\end{CodeChunk}
}

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on (Hartley,
Button, and Resta placed last).  In the first column, we see the
result from Australia (AUS) in which Hamilton came second, Vettel
first, Bottas third, and so on.  Notation used also includes six
different classes of no-score such as `Ret` for retired, `WD` for
withdrawn, and so on; a zero entry means "did not finish".  It is
straightforward to translate this table into a Plackett-Luce
log-likelihood function using the `hyper2` package; for simplicity we
will consider only the 10 top-ranked drivers\footnote{In the
Plackett-Luce likelihood function, the performance of lower-ranked
players can be informative about higher-ranked players' strengths.
For example, we see that Vettel retired twice (in Singapore and
Japan), so any player who placed in those venues will effectively
``steal'' strength from Vettel, and generally ``give'' it to Hamilton
or Bottas.}.  Although it has many terms, the first few are

```{r,echo=FALSE,print=FALSE}
library(hyper2,quietly=TRUE)
H <- ordertable2supp(keep(F1_table_2017,1:10))
head(H,5)
```

@@

```{r,echo=FALSE}
mH <- maxp(H)
print(mH)
pie(maxp(H)[c(1,6,2,7,3,8,4,9,5,10)])
```

We see the that the top three players, Hamilton, Vettel, Bottas, are
of roughly equal strength and indeed a likelihood ratio test
[`samep.test()`] does not reject the null of equality ($p=0.72$).

@@

The points scored by each driver give a ranking shown in the table;
but Plackett-Luce likelihoods also give a ranking that us unencumbered
by the arbitrary points given.  We can plot one ranking against the other:

```{r,label=likevspoints,cache=TRUE}
jj <- read.table("formula1_2017.txt",header=TRUE)
points <- jj[,21]
names(points) <- rownames(jj)

mL2017 <- maxp(H)

m <- mL2017   # maxp(H)
ox <- round(rank(-points)[1:10]) # negative because number 1 = largest number of points
oy <- rank(-m)      # negative because number 1 = largest strength
par(pty='s')        # square plot
plot(ox,oy,asp=1,pty='s',xlim=c(0,10.2),ylim=c(0,10.2),pch=16,
xlab="official order",ylab="my order",main='Formula 1, 2017 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray') }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
grid()
print(ox)
print(oy)
```

>>>>

\small
\begin{CodeChunk}
\begin{CodeOutput}
           AUS CHN BHR RUS ESP MON CAN AZE AUT GBR HUN BEL ITA SIN MAL JPN USA MEX BRA ABU
Hamilton     2   1   2   4   1   7   1   5   4   1   4   1   1   1   2   1   1   9   4   2
Vettel       1   2   1   2   2   1   4   4   2   7   1   2   3 Ret   4 Ret   2   4   1   3
Bottas       3   6   3   1 Ret   4   2   2   1   2   3   5   2   3   5   4   5   2   2   1
Raikkonen    4   5   4   3 Ret   2   7  14   5   3   2   4   5 Ret DNS   5   3   3   3   4
Ricciardo  Ret   4   5 Ret   3   3   3   1   3   5 Ret   3   4   2   3   3 Ret Ret   6 Ret
...
Hartley      0   0   0   0   0   0   0   0   0   0   0   0   0   0   0   0  13 Ret Ret  15
Button       0   0   0   0   0 Ret   0   0   0   0   0   0   0   0   0   0   0   0   0   0
Resta        0   0   0   0   0   0   0   0   0   0 Ret   0   0   0   0   0   0   0   0   0
\end{CodeOutput}
\end{CodeChunk}

```{r,echo=FALSE,print=FALSE}
top <- 11
```

Each row is a driver and each column (after the first) a venue.  We
see that Hamilton, the first row, came second in Australia, first in
China, second in Bahrain, fourth in Russia, and so on (Hartley,
Button, and Resta placed last).  In the first column, we see the
result from Australia (AUS) in which Hamilton came second, Vettel
first, Bottas third, and so on.  Notation used also includes six
different classes of no-score such as `Ret` for retired, `WD` for
withdrawn, and so on; a zero entry means "did not finish".  It is
straightforward to translate this table into a Plackett-Luce
log-likelihood function using the `hyper2` package; for simplicity we
will consider only the `r top` top-ranked drivers\footnote{In the
Plackett-Luce likelihood function, the performance of lower-ranked
players can be weakly informative about higher-ranked players'
strengths.  For example, we see that Vettel retired twice (in
Singapore and Japan), so any player who placed in those venues will
effectively ``steal'' strength from Vettel, and generally ``give'' it
to Hamilton or Bottas.}.  Although it has many terms, the first few
are

```{r,echo=FALSE,print=FALSE}
library(hyper2,quietly=TRUE)
top <- 11
H <- ordertable2supp(F1_table_2017[seq_len(top),])
head(H,5)
```

@@

```{r,echo=FALSE}
mH <- maxp(H)
print(mH)
pie(maxp(H)[c(1,6,11,2,7,3,8,4,9,5,10)])
```

We see the that the top two players, Hamilton and Vettel
of roughly equal strength and indeed a likelihood ratio test
[`samep.test()`] does not reject the null of equality ($p=0.09$).

```{r,cache=TRUE}
samep.test(H,1:2)
```

@@

The points scored by each driver give a ranking shown in the table;
but Plackett-Luce likelihoods also give a ranking that does not depend
on any arbitrary points system.  We can plot one ranking against the
other:

```{r,label=likevspoints,cache=TRUE,echo=FALSE}
jj <- read.table("formula1_2017.txt",header=TRUE)
points <- jj[,21]
names(points) <- rownames(jj)

mL2017 <- maxp(H)
m <- mL2017   # maxp(H)
ox <- round(rank(-points[seq_len(top)]))  # negative because number 1 = largest number of points
oy <- rank(-m)                            # negative because number 1 = largest strength
par(pty='s')        # square plot
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="official order",ylab="my order",main='Formula 1, 2017 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray') }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

From the above, we see that
